### 2017-02-27 paper/proposal.Rmd
Description of hugo's data

---
title: "Pollen trade-offs"
author: "Fernando, Hugo, Jason & Daniel"
date: "27 February 2017"
output:
  html_document: default
  word_document: default
subtitle: Proposal
---

# Introduction

Interspecific pollen deposition is detrimental to plant fitness. 
This is particularly true if it's the first deposition.
Using specialization: trade offs
Plants can minimise interspecific pollen deposition using four distinct mechanisms:

1. By preferentially flowering on spatial clusters 
2. By flowering in missmatching phenologies
3. By harnessing individual level insect preferences
4. By depositing and receiving pollen on/from specific insect body parts

I aim to use

# Methods

## Individual based model

## Hugo's data

Data collected over 3 years, each year with 5 samplings from November to March. 
For each sampling, 3 sites were visited, each of them with 4 fragments (two restored and two agricultural). 
In each fragment, 2 transects were performed. 
For each transect x sampling, there is data on flower abundance, insect abbundance (# insects caught), plant-insect visitation(?) and pollen carried per individual insect (to the species level). 
In addition, there is data on the amount of interspecific/conspecific pollen in flower stigmas for year #2. (From November to March as well?)
It might be possible to gather information on plant and animal phenophases and traits either from available data or databases / published literature.

# Timeline