### 2018-03-20 publication/manuscript.Rmd
Small changes to results

---
title: "The pollination trade-off"
author: "Fernando Cagua, Hugo Marrero, Jason Tylianakis, Daniel Stouffer"
bibliography: references.bib
csl: ecology-letters.csl
output: 
  bookdown::pdf_document2: 
    keep_tex: yes
    number_sections: false
toc: false
header-includes:
  - \usepackage{booktabs}
  - \usepackage{setspace}
  - \usepackage{lineno}
  - \usepackage{xr}
  - \externaldocument[S-]{supp_info}
---

\linenumbers
\doublespacing

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, 
                      warning = FALSE)
options(xtable.comment = FALSE, 
        xtable.caption.placement = 'bottom',
        xtable.size = 'small',
        booktabs = TRUE)
```

## Abstract

TBC

## Introduction

Pollination systems play a disproportionally important role in food production and maintenance of global biodiversity. 
In the context of animal pollination, the support to biodiversity is fundamentally achieved by virtue of intra-species facilition—the positive feedback loops that exist between plants that share pollinators or pollinators that share plants. 
Facilitation is able to promote species coexistence because it offsets the effects of direct competition for resources. 
Indeed, studies that are fundamental to our current knowledge of mutualism predict that an upper limit to biodiversity is achieved when the number of mutualistic partners is maximised in a community. 
And yet, fully connected natural pollination communities do not exist.

This is so, because sharing a mutualistic partner also has costs. 
The end result is that, when the cummulative costs have negative impact on fitnessis, species also start competeing for mutualistic partners. 
For the 88% of flowering plants that rely on animal pollination, there are two mechanisms trough which competition for pollinators might arrise. 
The first is competition for visits, where plants with more atractive flowers reduce the pollination service for those less atractive. 
The second is interspecific pollen transfer, where even receiving a visit might not translate into net facilitation. 
A focal plant might receive pollen from a different species, or conversely pollen from the focal plan might be lost to different species. 
Here we focus on this second mechanism.

The potential negative effects of competition for pollinators have been long recognised and widely documented. 
It is currently clear that they have the potential to drive the evolution of flower phenologies, morphologies and reproductive strategies as well as driving ecological diferentiation. 
Indeed there is considerable evidence demonstrating the effects that interspecific pollen transfer in two-species systems. 
However, these competitive interactions between species pairs do not occur in isolation. 
Instead, plants often ocurr in communities in which multiple competitive interactions operate simultaneusly and lead to emergent phenomena not observed at smaller scales.

Nevertheless, only a handful of studies have examined interspecific pollen transfer at the community level, and therefore we still understand little about how compeition for pollinators operates at this scale. 
All together, existing studies show that, as expected, the effective result of mutualistic interactions is, in most cases, one of facilitation. 
Importantly, they also show a large variability in this result—both across species and across communities. 
However, the driving factors behind this variability are currently not understood.

Understanding what makes species and communities different is important not only because of its ecological and evolutionary implications. 
It is also important because it is fundamentaly related to the way we understand and mathematically represent mutualistic systems. 
Some of the most widely used theoretical models of pollination account for the negative effects of mutualisim by adjusting the mutualistic benefit of an interaction so that it can be negatively affected by the number of mutualistic partners---the species degree. 
The extent to which species degree affects the benefit is called *mutualistic trade-off*. 
This approach, considerably simplifies the mathematical treatment of competition for pollinators. 
At the same time it is based on the intuitive rationale that a species that interacts with few species benefits strongly from each of them, whereas a species that interacts with a large number of species does so comparatively weakly. 
Although, the mutualistic trade-off has been shown to play a determining role on the stability and diversity of mutualistic communities, empirical evidence is scarce, to say the least.

Elucidating how the mechanisms and drivers of facilitation scales to the community level and its theoretical concequences has been elusive in part because of the large amounts of data required.
Previous studies focus on the final picture of interspecific pollen transfer that is obtained by examining the pollen deposited on flower stigmas. 
However, teasing appart the competition for pollinators also requires data that describes the structure of the interactions and the pollen flows that ocurr within it, while accounting for the traits of the species that integrate the community.
Here, we examine a comprehensive data set that includes information about pollen deposition, transfer, visitation, species abundance, and species traits across diverse communities in the Argentinean Pampas. 
Specifically, we first estimate the gain on pollen that can be attributed to the mutualistic interactions. 
Second, we investigate wether there is a relationship between this gain and a suite of variables that describe both the role of the plant species and the ecological context. 
Here, we pay special attention to the role of species degree because of its ramifications to theoretical ecology. 
At the same time, we examine wether the plant role and ecological context is best infered from dynamics occurring at the community level, or wether it can be well approximated using information available from all sites in which a species ocurr. 
Third, and finaly, we explore how plant species differ on the extent to which competition for pollinators shapes facilitation.

## Methods

### Data collection

We performed sampling in three locations, each located in a distinct phytogeograohic region of the Argentinean Pampas. 
The sampling locations are distributed across a gradient of precipitation with a humid temperate climate in the eastward Flooding Pampas and an drier as we move inland and westward towards the Pampean Grasslands.
Across all sites, annual mean temperatures range between 14.8 and 15.8 Celsius.

In each location, we sampled two restored and two agricultural fragments. 
However, due to the lack of available sites in the Flooding Pampas, it was only possible to sample one restored site.
We visited each fragment in November 2010 and February 2011. 
In each of these visits we gathered information related to abundance, plant-pollinator visitation, pollen transfer, and pollen deposition.

Abbundance was estimated using two 50 m randomly located transects in each fragment. In a first sampling we counted all units of floral attraction taht were found in a 2 m wide strip. In a second sampling we counted and collected all floral visitors while walking at a pace of 10 m per minute [@Marrero2014; @Memmott1999]. 
We constructed quantitative visitation networks using information about the floral visitors collected during abundance transects [@Marrero2014]. 
In addition, we also constructed qualitative visitation networks during independent two hour observations of floral visits across each fragment.
We estimated pollen transfer by examining the pollen loads present on the collected floral visitors [@Marrero2017]. 
Where the pollen count on an individual was estimated to be less than 2,000 grains, we identified every grain to the species level whenever possible and to pollen complexes when it was not. 
When the pollen count was above 2,000 grains, we clasified approx. 50% of pollen and total pollen counts were extrapollated [@Bosch2009]. 
If more than 10 pollen grains from a plant species were identified, we assumed that pollination service between the plant and the pollinator existed. 
Finally, we analised pollen deposition in a subset of the plant community [@Marrero2016]. 
This subset comprised between three and nine of the most common entomophilus species that were flowering during the samplimg period while ensuring that the chosen species covered a wide range on a specialization-generalization gradient. 
In the selected plants we removed all flowers except buds that were expected to go into inflorecense on the next day. 
A quarter of these buds were bagged to prevent animal pollination. 
Two days after inflorecense, we analysed the pollen grains in the flowers' pistils and classified them between conspecific and heterospecific pollen.

More details about the study sites, data collection, and laboratory protocols can be found in Marrero *et al.* (2014, 2016, and 2017).

### Data analysis

First, we explored wether animal pollination resulted in a quantitative increase in the pollination service.
To do that, we compared the amount of conspecific pollen on stigmas from flowers open to animal pollination and that from flowers that were bagged, where only autogamus self-pollination ocurred.
The difference between these two counts is not directly equivalent to the pollen deposited by animals because self-pollination is an adaptive trait [@Kalisz2003]. 
Nevertheless, this difference still provides valuable indication of what the animal contribution to reproductive output.
We performed this comparisons using a set of non parametric Mann-Whitney tests both at the community and the the species level (grouping species across communities).

Our main aim was to examine the extent to which the pollination service was influenced by the number of pollination partners while accounting for other metrics that describe the plant's pollination niche and community composition. 
To do so, we fitted four sets of linear mixed models [using the R package `nlme 3.1-131`, @nlme2017] which differed on whether the pollen deposited was conspecific or heterospecific and on wether independent variables were calculated across communities or independently within each community. 
In all set of models, our response variable was the difference between pollen density (pollen counts per stigma) between open and bagged flowers. 
All together the results from each model set—and the relationship between them—provide with two important pieces of information. 
First,  they provide indication of both the quantity and quality of animal pollination, and the potential trade-off between these two. 
Second, it allowed us to examine wether indeed community level information is necessary to explain the observed patterms or whether less granular data is sufficient.

All model sets were based on the same one hundred bootstrap resamples of the deposition data. 
We used bootstrap resampling because, as open and bagged flowers were not paired, this allowed us to directly model the *gain* in pollen density rather than density itself. 
This strategy, not only simplified the model specification and intepretation, but also allowed us to easily evaluate the uncertainity of the model results.  
In each replicate we randomly sampled with replacement a number of open flowers within each community-species combination and an equal number of closed flowers. 
We then log-transformed the pollen densities to improve normality and minimise the impact of outliers (Figure \@ref(S-fig:fig-deposition-logtransformed-or-not) and \@ref(S-fig:fig-deposition-gain-logtransformed-or-not)). 
Finally, we calculated the difference between the pollen density among treatmemts. 
This difference was the response variable in all our models.

As predictors, we included the species/degree (number of animal partners) as well as other factors that could influence the trade-off. 
Specifically, we included the plants' relative abundance and the temporal niche overlap. 
To calculate the plant's relative abbundance we simply aggregated the flower counts over the study period—either across communities of for each community depending on the model set. 
As with the deposition data, we found that log-transforming the abundance counts drastically improved normality (Figure \@ref(S-fig:fig-abundance-logtransformed-or-not)).
We then the log-transformed abundance counts for each month between November 2010 and March 2011 to calculate the temporal niche overlap between plants. 
Specifically, we used the Pianka's niche overlap index [@Pianka1973; `spaa 0.2.2` R package @Zhang2016] which calculates an overlap index for each pair of species—again, across and within  communities depending on the model set. 
To calculate an overlap metric per species, we simply averaged the pairwise indices for a species. 
To faciliate comparison across model estimates, we scaled all independent variables to have a mean of zero and a standard deviation of one.

In all set of models we first evaluated a set of candidate random effects which included random intercepts for plant species as well as random slopes for the relationship between degree and pollen gain. 
In addition we also tested random intercepts assuming that the plant species grouping was hierachically nested in either their community, the land use, and the locality. 
We selected the best random structure by comparing the median Akaike Information Criterion (AIC) of evaluated models fitted using restricted maximum likelyhood. 
After determining the best random structure, we then focused on the model's fixed structure where we fitted the fixed effect coefficients using maximum likelyhood. 
Because our purpose was to perform inter-model comparisons across multiple data resamples, we did not perform model selection. 
Instead, in all model sets, we obtained the estimates for all the fixed effects we considered and then evaluated their relevance by examining their distribution.

We compared the performance of the within and across communities model sets using the root-mean-square error normalised by the interquartile range (NRMSE). 
We tested the significance of this difference using a two sample paired Wilcoxon signed rank test for conspecific and heterospecific pollen models. 
We chosed NRMSE because it provides an easily interpretable metric related to the quality of the predictions while being farily insensitive to sample size. 
This inensitivity was desirable because the across communities models had often slightly larger sample sizes owing to the fact of some species not having all data necessary to calculate the independent variables at the community level.

Finally, we focused on how the relationship between conspecific and heterospecific pollen gain. 
First we examined wether the pollination trade-off was related to our explanatory variables. 
To do so, we harnessed the results from the conspecific and heterospecific pollen models that performed best (across or within communities). 
To examine the relationship between these model sets, we sampled 100 estimates from the distribution of coefficients in each set. 
For each of the samples, we then calculated the difference between the gain in conspecific and heterospecific pollen that would be predicted assuming mean values for the random intercepts and the other predictors.

Second, we calulated the the extent of facilitation. 
For this purpose, as in @Tur2016a, we used the relationship between heterospecific and conspecific pollen. 
We calculated an overal relationship by sampling the predictions from 100 conspecific/heterospecific model pairs and estimated the relationship by fitting an standardised major axis (SMA) to each pair [we used the R package `smatr 3.4-3`, @Warton2012a]. 
We followed a similar procedure to calculate a relationship per species but instead of fitting a SMA per model pair, we did so across the 100 models for each species independently.

## Results

```{r differences-in-self-pollination}
loadd(significant_gain_global, significant_gain_site)

sgg <- glance_table(significant_gain_global)
prop_sig_global <- sum(sgg$p.value < 0.05) /
  length(significant_gain_global)
sgs <- glance_table(significant_gain_site)
prop_sig_site <- sum(sgs$p.value < 0.05) /
  length(significant_gain_site)

consistency <- inner_join(sgg, sgs, by = 'recipient') %>% 
  mutate_at(.vars = c('p.value.x', 'p.value.y'), 
            .funs = function(x) x < 0.05) %>%
  mutate(consistent = p.value.x == p.value.y) %>%
  group_by(recipient) %>%
  summarise(p.value.x = mean(p.value.x), 
            consistent = all(consistent),
            n_sites = n_distinct(site_name)) %>% 
  filter(n_sites > 1)
prop_consistency <- sum(consistency$consistent)/
  nrow(consistency)
```

Over all study sites, we found that `r round(prop_sig_global*100)`% of species observed a significant increasse (at the $\alpha = 0.05$ level) in pollen density when comparing open and bagged flowers (Table \@ref(S-tab:table-species-gain-global)). 
When looking at the differences within sites, we found that only `r round(prop_consistency*100)`% of the species that were present in more than one site had the same response across community (Table \@ref(S-tab:table-species-gain-site)). 
Despite these differences, the community had a comparatively small effect on wether animal pollination increased the pollen density or not (Table \@ref(S-tab:table-variation-across-sites)).

We found that the random structure that was best supported the data was consistently the one that included plant species nested in community as the grouping factors (Table \@ref(S-tab:table-aic-random-models)). 
When comparing the performance of the models constructed at the community or study-wide level we found that the gain in conspecific pollen was better approximated by the model sets using predictors calculated across communities (Table \@ref(S-tab:table-global-vs-community-models)). 
This was so despite the fact that the model predictors calculated at the community level were highly correlated to those at the study wide level (Figure \@ref(S-fig:fig-global-community-correlations)). 
Nevertheless, results from the within community models are similar both qualitatively and quantitatively (with the exception of phenology overlap, but that might change once I use a better metric, Figure \@ref(S-fig:fig-model-results)).

We found that the gain in conspecific pollen is negatively affected by the plants relative abbundance and to a smaller extent by the species' degree (Figure \@ref(fig:fig-model-results-global). 
The gain in heterospecific pollen was instead positively affected by degree and the plant's phenology overlap, but was also strongly negatively affected by the plant's relative abbundance.

```{r fig-model-results-global, fig.height = 2, fig.cap = "Distribution of effect estimates for models of conspecific and heterospecific pollen density gain. Results are shown for models with explanatory var"}
loadd(tidied_fixed, sites)

tidied_fixed %>%
  filter(term != "(Intercept)" , 
         var_trans == 'log',
         scale == "global"
         ) %>%
  group_by(pollen_category,scale, model, var_trans, term) %>%
  summarise(estimate = first(estimate)) %>%
  group_by() %>%
  humanize() %>%
  ggplot(aes(x = estimate, 
             colour = pollen_category, 
             fill = pollen_category)) +
  geom_vline(xintercept = 0, linetype = 2, size = 0.25) +
  stat_density(geom = 'line',
               # alpha = 0.15,
               position = position_identity(),
               draw_quantiles = c(0.5, 0.95, 0.05), 
               trim = T, 
               scale = "count",
               bins = 15) +
  pub_theme() +
  facet_grid( ~ term, scales = "free_x", space = "free_x") +
  scale_color_manual(values = colour_scale_values) +
  scale_fill_manual(values = colour_scale_values) +
  scale_x_continuous(breaks = seq(-2,2, by = 0.25)) + 
  theme(legend.position = "top") +
  labs(colour = "", fill = "", x = "estimated effect on pollen gain")
```

The gain in conspecific pollen is positively correlated with the gain in heterospecific pollen. 
The slope of the relationship is such that it is more likely that the gain on heterospecific pollen increases faster than the gain on conspecific pollen (Figure \@ref(fig:fig-trade-off-predictions)A). 
In addition, we found that from all the explanatory variables examined, only the plant's relative abbundance has a positive effect on the proportion of conspecific pollen gained (Figure \@ref(fig:fig-trade-off-predictions)B). 
For most of the parameter space plants obtain a larger proportion of heteropspecific than conspecific pollen. 
However, plants can obtain a larger proportion of conspecific pollen when they have high abundance and to a smaller extent when they have a low degree. 
When examining the extent of facilitation at the species level, we found that XXX (Figure \@ref(fig:fig-trade-off-predictions)B).

```{r fig-trade-off-predictions, fig.height= 2.5, fig.cap = "The interplay between the gain on heterospecific and conspecific pollen. (A) Overall and species-level relationship. Grey lines indicate the relationship in each of the samples, the median is indicated with a black solid line. Dotted lines show the relationship for each species. Invasive species are shown in red. (B) The proportion of conspecific pollen relative to the total amount of pollen gained trough animal pollination and its relationship to our set of explanatory variables."}
loadd(trade_off_predictions, model_linear_fits, tidied_fixed)

get_pred_range <- function(x){
  tidied_fixed %>% 
    filter(pollen_category == x, var_trans == "log", scale == "global") %$% 
    estimate %>% range()
}

ab_lines <- model_linear_fits %>%
  filter(var_trans == "log",
         scale == "global") %>%
  spread(sma_parameter, value)

p1 <- data_frame(x = get_pred_range("heterospecific"),
                 y = get_pred_range("conspecific")) %>%
  ggplot(aes(x, y)) +
  geom_abline(data = ab_lines, 
              aes(slope = slope, intercept = elevation), 
              size = 0.1, linetype = 1, colour = "grey", 
              alpha = 0.5) +
  geom_abline(data = summarise_if(ab_lines, is.numeric, median), 
              aes(slope = slope, intercept = elevation), 
              size = 0.5, linetype = 1, colour = "black") +
  geom_abline(slope = 1, intercept = 0, size = 0.25, linetype = 2) +
  geom_point(alpha = 0) +
  coord_equal(xlim = get_pred_range("heterospecific"),
              ylim = get_pred_range("conspecific")) +
  pub_theme() +
  labs(x = "gain in heterospecific pollen", 
       y = "gain in conspecific pollen")

p2 <- trade_off_predictions %>%
  mutate_at(c("conspecific", "heterospecific"), exp) %>%
  mutate(gain = conspecific / (conspecific + heterospecific)) %>%
  filter(!(value > 2.2 & term == "k")) %>%
  humanize() %>%
  ggplot(aes(x = value, y = gain, colour = term)) +
  geom_hline(yintercept = 0.5, size = 0.25, linetype = 2) + 
  geom_point(shape = 21, alpha = 0.5, size = 1) +
  geom_smooth(aes(fill = term), 
              se = F, size = 0.5, method = "glm", 
              method.args = list(family = "binomial"), alpha = 0.1)+
  pub_theme() +
  scale_color_manual(values = colour_scale_values) +  
  scale_fill_manual(values = colour_scale_values) +
  scale_y_continuous(labels = scales::percent) +
  theme(legend.position = "top") +
  labs(colour = "", fill = "", linetype = "", 
       y = "proportion conspecific pollen gain", 
       x = "scaled variable value")

cowplot::plot_grid(p1, p2, 
                   align = "h", axis = "tb", ncol = 2, labels = LETTERS[1:2], 
                   label_size = 8, rel_widths = c(0.35, 0.65))
```

## Discussion

This is, the data is best modelled when ignoring the hierarchical nature of the experiment (communities nested in agricultural or restored fragments nested in localities).
Similarly, the relationship between degre and pollen gain does not seem to be species or community dependent.

Relative small difference between conspecific animal and self pollination is conservative because crossing is better than cloning ()

## References