### 2019-03-17 paper/supp-info.Rmd
Remove unevaluated chunks from supp. info

---
title: "The pollination trade-off"
subtitle: "Supplementary information"
author: "Fernando Cagua, Hugo Marrero, Jason Tylianakis, Daniel Stouffer"
# bibliography: references.bib
csl: ecology-letters.csl
documentclass: book
nocite: | 
  @Xu2003, @Nakagawa2013
output: 
  bookdown::pdf_document2: 
    keep_tex: yes
    toc: false
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
editor_options: 
  chunk_output_type: console
---

<!-- To add an S before each figure and table -->
\renewcommand\thefigure{S\arabic{figure}}    
\setcounter{figure}{0}

\renewcommand\thetable{S\arabic{table}}    
\setcounter{table}{0}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, 
                      warning = FALSE)
options(xtable.comment = FALSE, 
        xtable.caption.placement = 'bottom',
        xtable.size = 'small',
        booktabs = TRUE, 
        knitr.kable.NA = "-")
```

```{r table-facilitation-models-general, results='asis'}
drake::loadd(facilitation_models)

broom::tidy(facilitation_models, conf.int = F) %>%
  dplyr::select(-p.value, -group) %>%
  dplyr::mutate(term = dplyr::case_when(
    # term == "(Intercept)" ~ "Intercept",
    term == "sd_(Intercept).plant_name:site_name" ~ "S.D. random intercept", 
    term == "sd_heterospecific.plant_name:site_name" ~ "S.D. random slope", 
    TRUE ~ term
  )) %>%
  dplyr::filter(term != "cor_(Intercept).heterospecific.plant_name:site_name") %>%
  knitr::kable("latex", 
               caption = "Summary of the model used to analyse the relationship between heterospecific and conspecific pollen", 
               digits = 3,
               booktabs = T,
               col.names = c("predictor", "estimate" ,"S.E.", "z-value")) %>%
  kableExtra::kable_styling(font_size = 8) %>%
  kableExtra::group_rows(index = c("fixed component" = 2, 
                                   "random component (species:community)" = 2)) 
```

```{r table-facilitation-models-random, results='asis'}
drake::loadd(facilitation_random_effects)
drake::loadd(sites)

tbl_frame <- facilitation_random_effects %>%
 dplyr::inner_join(sites, by = "site_name") %>%
  dplyr::mutate(site_name = paste(locality_name, 
                                  land_use, 
                                  fragment, 
                                  sep = ' - ')) %>%
  dplyr::mutate_if(is.character, knitr:::escape_latex) %>%
  dplyr::mutate(plant_name = paste0("\\textit{", plant_name, "}")) %>%
  dplyr::select(plant_name, site_name, heterospecific, error)

tbl_frame %>%
  knitr::kable("latex", 
               digits = 4, 
               col.names = c("species name", "community", "slope", "S.E."), 
               caption = "Slope of the relationship between heterospecific and conspecific pollen for each species in their community (fixed effect + conditional effect). Community names are constructed by location - agricultural/restored - fragment number.", 
               booktabs = T, 
               linesep = "", 
               escape = F) %>%
  kableExtra::kable_styling(font_size = 8) 
```

```{r table-open-bagged-models-general, results = "asis"}
drake::loadd(open_bagged_model)

broom::tidy(open_bagged_model, conf.int = F) %>%
  dplyr::select(-p.value, -group) %>%
  dplyr::mutate(term = dplyr::case_when(
    # term == "(Intercept)" ~ "Intercept (bagged flower)",
    term == "treatmentopen" ~ "treatment (unbagged flower)",
    term == "sd_(Intercept).plant_name:site_name" ~ "S.D. random intercept", 
    term == "sd_treatmentopen.plant_name:site_name" ~ "S.D. random slope", 
    TRUE ~ term
  )) %>%
  dplyr::filter(term != "cor_(Intercept).treatmentopen.plant_name:site_name") %>%
  knitr::kable("latex", 
               caption = "Summary of the model used to analyse the relationship between conspecific pollen deposited in bagged and unbagged flowers (open to animal pollination).", 
               digits = 3,
               booktabs = T,
               col.names = c("predictor", "estimate" ,"S.E.", "z-value")) %>%
  kableExtra::kable_styling(font_size = 8) %>%
  kableExtra::group_rows(index = c("fixed component" = 2, 
                                   "random component (species:community)" = 2)) 
```

```{r table-open-bagged-models-random, results='asis'}
drake::loadd(con_con_plot_df)

con_con_plot_df %>%
 dplyr::inner_join(sites, by = "site_name") %>%
  dplyr::mutate(site_name = paste(locality_name, 
                                  land_use, 
                                  fragment, 
                                  sep = ' - ')) %>%
  dplyr::mutate_if(is.numeric, round, digits = 0) %>%
  dplyr::mutate(open_ci = paste0("[", open_lower, ", ", open_upper, "]"), 
                closed_ci = paste0("[", closed_lower, ", ", closed_upper, "]")) %>%
  dplyr::mutate_if(is.character, knitr:::escape_latex) %>%
  dplyr::mutate(plant_name = paste0("\\textit{", plant_name, "}")) %>%
  dplyr::select(plant_name, site_name, closed_mid, closed_ci, open_mid, open_ci) %>%
  knitr::kable("latex", 
               col.names = c("species name", "community", "mean", "C.I.", "mean", "C.I."), 
               caption = "Amount of conspecific pollen deposited  in bagged and unbagged flowers (open to animal pollination) for each species in their community. Community names are constructed by location - agricultural/restored - fragment number.", 
               booktabs = TRUE, 
               linesep = "", 
               align = "llrlrl", 
               escape = FALSE) %>%
  kableExtra::kable_styling(font_size = 8) %>%
  kableExtra::add_header_above(c(" " = 2, "bagged" = 2, "unbagged" = 2))
```

```{r table-R2, results="asis"}
drake::loadd(r2_values)

r2_values %>%
  dplyr::select(r2c_mean, r2c_min, r2c_max, r2m_mean, r2m_min, r2m_max) %>%
  knitr::kable("latex", 
               col.names = c(rep(c("mean", "min", "max"), times = 2)), 
               caption = "Coefficient of determination $R^2$ of the most parsimonious pollen deposition models (those with the lowest AICc). The marginal coefficient of determination describes the proportion of variance explained by just the fixed effects.", 
               booktabs = TRUE) %>%
  kableExtra::kable_styling(font_size = 8) %>%
  kableExtra::add_header_above(c("conditional $R^2_{(c)}$" = 3,
                                 "marginal $R^2_{(m)}$" = 3), 
                               escape = FALSE) %>%
  kableExtra::group_rows(index = c("conspecific pollen" = 1, 
                                   "heterospecific pollen" = 1))
```

```{r table-aic-random-models, results= 'asis'}
drake::loadd(glanced_random, random_effects)

summ <- glanced_random %>%
  dplyr::filter(var_trans == "log", 
                scale == "community") %>%
  dplyr::group_by(pollen_category, scale, var_trans, model) %>%
  dplyr::mutate(delta_AIC = AIC - min(AIC)) %>%
  dplyr::group_by(pollen_category, scale, var_trans, random_effect) %>%
  dplyr::summarise(AIC_50 = median(delta_AIC), 
            AIC_05 = quantile(delta_AIC, 0.05),
            AIC_95 = quantile(delta_AIC, 0.95), 
            n = n()) %>% 
  dplyr::group_by() %>%
  dplyr::select(-var_trans) %>%
  dplyr::right_join(expand.grid(pollen_category = unique(.$pollen_category),
                        scale = unique(.$scale),
                        random_effect = unique(.$random_effect))) %>% 
  dplyr::mutate_if(is.numeric, round, digits = 1) %>% 
  dplyr::mutate(delta_AIC = AIC_50, 
                delta_AIC_CI = paste0("[", AIC_05, ", ", AIC_95, "]")) %>%
  dplyr::group_by() %>%
  humanize(random_effects = random_effects) %>%
  dplyr::filter(pollen_category %in% c("conspecific (absolute)", "heterospecific")) %>%
  dplyr::mutate(
    pollen_category = dplyr::if_else(
      pollen_category == "conspecific (absolute)", 
      "conspecific pollen", "heterospecific pollen"), 
    random_name = knitr:::escape_latex(random_name)) %>%
  dplyr::arrange(pollen_category, delta_AIC) %>%
  dplyr::rename(`pollen type` = pollen_category,
         `random structure` = random_name)

summ %>%
  dplyr::select(`random structure`, delta_AIC, delta_AIC_CI) %>%
  kableExtra::kable("latex", 
               caption = paste('Comparison of the two random structures we considered for the models of conspecific and heterospecific pollen deposition. The table shows median $\\Delta$AIC values of', dplyr::n_distinct(glanced_random$model), "bootsrap resamples of the data. The 5th and 95th percentile are shown inside square brackets. Communities are defined by individual fragments but ignore the hierarchical arrangement of sampling sites."),
               col.names = c("random structure", "median", "C.I."),
               booktabs = TRUE, 
               escape = F) %>%
  kableExtra::kable_styling(font_size = 8) %>%
  kableExtra::group_rows(index = c("conspecific pollen" = 2, 
                                   "heterospecific pollen" = 2)) %>%
  kableExtra::add_header_above(c(" " = 1, "$\\\\Delta$AIC" = 2), escape = F)
```

```{r table-aic-fixed-models, results= 'asis', eval=T}
drake::loadd(model_formula_ranking)

model_formula_ranking$by_model_set %>% 
  dplyr::filter(scale == "community") %>%
  dplyr::mutate_if(is.numeric, round, digits = 1) %>%
  dplyr::mutate(delta_AIC_CI = paste0(" [", delta_AIC_quantile_05,
                                      " ,", 
                                      delta_AIC_quantile_95,
                                      "]")) %>% 
  dplyr::group_by() %>%
  humanize(formula_long = TRUE) %>%
  dplyr::filter(pollen_category %in% c("conspecific (absolute)", "heterospecific")) %>%
  dplyr::mutate(
    pollen_category = dplyr::if_else(
      pollen_category == "conspecific (absolute)", 
      "conspecific pollen", "heterospecific pollen")) %>% 
  dplyr::mutate(fixed_formula = paste("~", fixed_formula)) %>%
  dplyr::arrange(pollen_category, delta_AIC_rank) %>%
  dplyr::select(fixed_formula, delta_AIC, delta_AIC_CI) %>%
  dplyr::rename(`fixed structure` = fixed_formula) %>% 
  knitr::kable("latex", 
               col.names = c("fixed structure", "median", "C.I."),
               caption = paste('Comparison of the different fixed structures we consideredfor the models of conspecific and heterospecific pollen deposition. The table shows median $\\Delta$AIC values of', dplyr::n_distinct(glanced_random$model), "bootsrap resamples of the data. The 5th and 95th percentile are shown inside square brackets."),
               booktabs = TRUE) %>%
  kableExtra::kable_styling(font_size = 8) %>%
  kableExtra::group_rows(index = c("conspecific pollen" = 15, 
                                   "heterospecific pollen" = 15)) %>%
  # kableExtra::collapse_rows(columns = 1, latex_hline = "major", valign = "middle") %>%
  kableExtra::add_header_above(c(" " = 1, "$\\\\Delta$AIC" = 2), escape = F)
```

```{r fig-correlation, fig.height = 5, fig.cap = "Corelation between the explanatory variables included in the statistical models.", eval=T}
drake::readd(fig_correlation)
```

```{r fig-model-results, fig.height = 5, fig.cap = "Distribution of effect estimates for models of conspecific and heterospecific pollen density gain. Model formulas have been abbreviated: *a* for abundance, *k* for the number of shared polinators, *p* for the visit effectiveness, and *t* for trait originality. Only candidate formulas with a $\\Delta AICc < 4$ for either conspecific or heterospecific pollen are shown. Models candidates are arranged in decreassing order of support.", eval=T}
drake::readd(fig_all_model_results)
```

```{r fig-global-community-correlations, fig.height = 2, fig.cap = "Relationship between the model's independent variables calculated at the community level and at the study-wide level.", eval=F}
drake::readd(fig_community_global_scatter)
```