### 2019-04-09 paper/cover-letter.tex
Add draft cover letter

%-----------------------------------------------------------------------------%
% Letter class
\documentclass[a4paper, 10pt]{letter}
% Name of sender
\name{Fernando Cagua}
% Signature of sender
\signature{Fernando Cagua}
% Address of sender
\address
{
    Fernando Cagua,\\
    University of Canterbury, \\
    Private Bag 4800, Christchurch 8041, \\
    New Zealand.
}
%-----------------------------------------------------------------------------%
\begin{document}
% Name and address of receiver
\begin{letter}
{
    Tim Coulson,\\
    Editor-in-chief,\\
    Ecology Letters
}
% Opening statement
\opening{Dear Kathleen,}
% Letter body
I am aware that you are pretty miffed at me for being the cause of the demise of your bookstore. You have to believe in me, that was not my intention at all.
I hope we can talk about this over coffee. How about we meet this Friday evening at the corner cafe? Let me know.
% Closing statement
\closing{Yours truly,}
\end{letter}
\end{document}
%-----------------------------------------------------------------------------%