### 2019-06-13 paper/draft-info.Rmd
Update date

---
title: "Draft information"
subtitle: "version: draft-7.2"
date: "2019-06-14"
# author: "Fernando Cagua, Hugo Marrero, Jason Tylianakis, Daniel Stouffer"
# bibliography: bibliography.bib
# csl: nature.csl
output:
  bookdown::pdf_document2:
    # base_format: rticles::peerj_article
    keep_tex: yes
    number_sections: false
toc: false
classoption: a4paper
geometry: 
  - textwidth=33em
  - textheight=48\baselineskip
# classoption: twocolumn
header-includes:
  - \usepackage{booktabs}
  - \usepackage{setspace}
  - \usepackage{lineno}
  - \usepackage{xr}
  - \usepackage[utf8]{inputenc}
  - \newcommand{\R}[1]{\label{#1}\linelabel{#1}}
  - \newcommand{\lr}[1]{page~\pageref{#1}, line~\lineref{#1}}
  - \externaldocument[M-]{manuscript}
---

\onehalfspacing

## Changes since the last draft

* Adjusted cover letter to submit to Nature Ecology and Evolution