### 2019-08-16 paper/drafts-and-revisions/2019-07-05/review-response-notes.md
Some notes on how to address the reviews

# Review response notes

Each section represents a major issue that needs to be addressed in the manuscript

## Heterospecific-conspecific slope and competition/facilitation

The fact that a negative slope might indicate competition and a posititve slope between conspecific and heterospecific pollen indicate facilitation is probably the most contentious point of both reviews and the editor. 
Comments indicate issues related to the assumptions, methods, and conclusions. 
The reference to Tur et al. 2016, where this method was proposed was not enough to counter the reviewer's concerns.

**Way forward:** Given the amount and range of the critisism and that there is not many ways to address it, I think the best approach is to remove this section altogether. It might require a change on the paper focus from less competition/facilitation to a more pollen loads at the community level paper.

## Pollen deposition and Male-female fitness

The main critisism is that heterospecific pollen loads in the stigmas may not be detrimental to female fitness in some species. For the male fitness, pollen found in a species might not indicate losses to male fitness of another species. 
This is all true, but we can argue that heterospecific pollen deposition and conspecific pollen loss is neutral at best. 
Might not be bad for some species but doesn't help reproduction in any way.

**Way forward:** Emphasise non-positive effects. Removing first hypothesis from the paper might help here as well. Discuss that weakness of sharing pollinator effect might be due to this.

## Self pollen loads

Associate editor mentions that self-pollen might be as detrimental as heterospecific pollen and there is no way to distinguish self-pollen from other conspecific pollen.

**Way forward:** Put self-pollination back into the modelling framework

## Confounding 'evolutionary specialization' and 'functional originality'

Not sure what is meant by the reviewer here. Other than functional originality is dependent on the community while evolutionary specialization is intrinsic. Problem is lines 81-84.

**Way forward:** Unsure. Reduce the emphasis on community or remove from intro.

## Chosen traits

Reviewer states that traits should be properly justified. Specifically commented that life-history traits are not relevant in the context of pollination as that are subject to a different set of selective pressures.

**Way forward:** Remove traits or use phylogeny instead. Phylogeny might not solve the main issue though as it encompasses an even broader set of selective pressures.

## Niche flexibility test

Reviewer states that this hypothesis is best tested

## Study sites

## Species description

## Model specification

## Discussion claims