### 2019-08-16 paper/drafts-and-revisions/2019-07-05/review-response-notes.md
Add extra items to response notes

# Review response notes

Each section represents a significant issue that we need to address in the manuscript.

## Heterospecific-conspecific slope and competition/facilitation

The fact that a negative slope might indicate competition and a positive slope between conspecific and heterospecific pollen indicate facilitation is probably the most contentious point of both reviews and the editor. 
Comments indicate issues related to the assumptions, methods, and conclusions. 
The reference to Tur et al. 2016, where this method was first proposed was not enough to counter the reviewer's concerns.

**Way forward:** Given the amount and range of the criticism and that there are not many ways to address it, I think the best approach is to remove this section altogether. It might require a change on the paper focus from less competition/facilitation to more pollen loads at the community level paper.

## Pollen deposition and Male-female fitness

The main criticism is that heterospecific pollen loads in the stigmas may not be detrimental to female fitness in some species. For the male fitness, pollen found in a species might not indicate losses to male fitness of another species. 
These observations are all true, but we can argue that heterospecific pollen deposition and conspecific pollen loss is neutral at best. 
Might not be bad for some species but doesn't help reproduction in any way.

**Way forward:** Emphasise non-positive effects. Removing the first hypothesis from the paper might help here as well. Discuss that weakness of sharing pollinator effect might be due to this.

## Self pollen loads

The associate editor mentions that self-pollen might be as detrimental as heterospecific pollen and there is no way to distinguish self-pollen from other conspecific pollen.

**Way forward:** Put self-pollination back into the modelling framework

## Confounding 'evolutionary specialization' and 'functional originality'

Not sure what is meant by the reviewer here. Other than functional originality is dependent on the community while evolutionary specialization is intrinsic. Problem is lines 81-84.

**Way forward:** Unsure. Reduce the emphasis on community or remove from the intro.

## Chosen traits

The reviewer states that traits should be adequately justified. Specifically commented that life-history traits are not relevant in the context of pollination as that are subject to a different set of selective pressures.

**Way forward:** Remove traits or use phylogeny instead. Phylogeny might not solve the main issue though, as it encompasses an even broader set of selective pressures.

## Niche flexibility test

The reviewer states that we can test this hypothesis better by comparing variation within communities vs. across communities. That comment is weird cause our variables are established at the community level and hence there is no within community variability in our dataset.

**Way forward:** Cannot think of anything to do here, except try to explain the methods better?

## Study sites

The reviewer asks for more info about study sites. He remarks on why we didn't take into account the difference between agricultural and restored sites.

**Way forward:** Include more info about the whole methods, even though it is published. Perhaps we should put this in the supporting info so that the paper doesn't get too long?

## Species description

The reviewer asks for more information about the study plans, abundances, etc.

## Model specification

The reviewer suggests modelling ratio conspecific and heterospecific directly. Not sure about this, though.

**Way forward:** This seemed to be a minor complain though. Might get away with the same model? Another possibility is to use a single model for the whole thing

## Discussion claims

The reviewer suggests many of claims in the discussion are unsubstantiated.

**Way forward:** Just more nuance in the discussion. I think the reviewer had a pretty negative approach by then and was being a bit nitpicky.