### 2019-11-23 paper/manuscript.Rmd
Remove competition/niche lingo from results

---
title: "Community context determines competition vs. facilitation trade-offs in pollination systems"
author: 
  - "E. Fernando Cagua^1^ ([efc29@uclive.ac.nz](mailto:efc29@uclive.ac.nz))"
  - "Hugo J. Marrero^2^ ([hugomarrero@gmail.com](hugomarrero@gmail.com))"
  - "Jason M. Tylianakis^1^ ([jason.tylianakis@canterbury.ac.nz](jason.tylianakis@canterbury.ac.nz))"
  - "Daniel B. Stouffer^1^ ([daniel.stouffer@canterbury.ac.nz](daniel.stouffer@canterbury.ac.nz))"
bibliography: bibliography.bib
csl: proceedings-of-the-royal-society-b.csl
output: #word_document
  bookdown::pdf_document2:
    # base_format: rticles::peerj_article
    keep_tex: yes
    number_sections: false
toc: false
fontsize: 11pt
classoption: a4paper
geometry: 
  - textwidth=33em
  - textheight=48\baselineskip
header-includes:
  - \usepackage{booktabs}
  - \usepackage{setspace}
  - \usepackage{lineno}
  - \usepackage{xr}
  - \usepackage[utf8]{inputenc}
  - \newcommand{\R}[1]{\label{#1}\linelabel{#1}}
  - \newcommand{\lr}[1]{page~\pageref{#1}, line~\lineref{#1}}
  - \externaldocument[S-]{supp-info}
  - \usepackage[export]{adjustbox}
  # - \usepackage[nomarkers]{endfloat}
---

\linenumbers
\doublespacing

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, 
                      warning = FALSE, dpi = 300)

loadd(abs_wordcount, msc_wordcount, n_displays, n_references, keywords, references)
```

^1^ Centre for Integrative Ecology, School of Biological Sciences, University of Canterbury, Private Bag 4800, Christchurch 8041, New Zealand

^2^ Centro de Recursos Naturales Renovables de las Zonas Semiáridas, CONICET, Camino de la Carrindanga Km. 7, 8000 Bahía Blanca, Argentina

<!-- **Short title:** Competition/facilitation in pollination -->

**Content type:** Research article.

**Number of words:** `r abs_wordcount$n_words_stri` in abstract; `r format(round(msc_wordcount$n_words_stri), big.mark = ",")` in main text (including Methods section).

<!-- **Number of displays:** `r n_displays$figures` figures; `r n_displays$tables` tables; `r n_displays$text_boxes` text boxes. -->

<!-- **Number of references:** `r n_references` -->

**Author for correspondence:** E. Fernando Cagua (+64 20 4026 8153).

<!-- **Data accessibility:** All data and code necessary to reproduce the results has been included in the Electronic Supplementary Materials and will be publicly available in appropriate databases upon acceptance.. -->

<!-- **Author contributions:** EFC wrote the manuscript and performed all analyses with input from all authors; HJM collected the data; all authors contributed to the study design.  -->

<!-- **Competing interests:** The authors have declared that no competing interests exist. -->

<!-- The authors do not wish to opt-in to post the submission as a preprint on bioRxiv -->

\clearpage

# Abstract

`r paste(drake::readd(abstract), collapse = "")`

**Keywords:** `r keywords`

\clearpage

# Introduction

Animal pollination plays a disproportionally important role in food production and maintenance of global biodiversity [@bascompte_plant-animal_2007; @klein_importance_2007; @ollerton_how_2011]. 
At a pairwise level, the mutually beneficial relationship between plants and pollinators underpins pollination. 
At a community level, sometimes involving hundreds of species, both plant and pollinator species are connected in a myriad of indirect connections when pollination partners are shared. 
These indirect connections can dramatically alter the quality of the pollination service that plants receive because they determine how conspecific and heterospecific pollen is transferred across the community [@morales_interspecific_2008]. 
Generally speaking, there is a trade-off between the benefits gained from maximising the number of partners and the costs of sharing them with other plant species [@waser_interspecific_1978].
However, due to the large number of factors that operate at the community level, we do not understand how sharing pollinators affect the pollination service beyond systems with more than a handful of species. 
Here we investigate how pollinator sharing affects pollen transfer *in natural communities* and how it compares to other factors known to play a role in community dynamics like abundance, traits, and visitation patterns.

There are two main mechanisms through which sharing pollinators can affect plant fertilisation [@morales_interspecific_2008]. 
The first is by changes in intraspecific pollen transfer.
Interspecific pollen transfer happens, for example, when plants with more attractive flowers might reduce the number of visits to those less of attractive neighbouring plants, and hence reduce the amount of *conspecific pollen* deposited by animals [@yang_pollinator_2011]. 
The second is via interspecific pollen transfer. 
In that case, even receiving a visit might not necessarily translate into fertilisation [@campbell_mechanism_1985] because a focal plant might receive *heterospecific pollen* or because pollen from the focal plant might be lost to different species. 
Naturally, the precise effects on female or male plant fitness of conspecific and heterospecific pollen deposition depend on the species involved (and are unknown for many plant species). 
However, even for species well adapted to pollinator sharing, receiving foreign pollen on stigmas or losing pollen to foreign stigmas, is neutral at best. 
Indeed, there is substantial evidence supporting the idea that heterospecific pollen deposition can be detrimental to seed production and plant fitness [@ashman_toward_2013; @arceo-gomez_invasion_2016].

All together, provided pollen is viable and compatible [@ramsey_pollen_2000; @de_jong_plant_1992; @dafni_pollen_2000], the higher the quantity of conspecific pollen and its purity (relative to heterospecific pollen), the better the pollination service received by the focal plant. 
As such, measuring conspecific and heterospecific pollen deposition provides a good indication of the potential levels of facilitation and competition a plant population might experience.

By definition, intra- and interspecific pollen transfer, occur at the community scale.
However, with few exceptions [@tur_evaluating_2016; @aizen_reproductive_2010], most of what we know about pollen transfer and its relationship with key ecological factors are based on studies with two plant species. 
That is partly so because, at the community scale, the factors that determine the patterns of pollen deposition are tightly intertwined, operate simultaneously, and may lead to emergent phenomena not observed at smaller scales [@flanagan_effects_2011].
For instance, recent empirical evidence suggests that plants with flowering traits that are "original" relative to others in the community generally have fewer interaction partners [@coux_linking_2016].

This evidence aligns with the notion that a species that interacts with few species does so strongly with each of them whereas a species that interacts with a large number of species does so comparatively weakly [@thebault_does_2008; @bascompte_asymmetric_2006; @vazquez_species_2007].
If evolutionary specialisation occurs by changing traits to focus on fewer but better partners [@caruso_competition_2000], we should expect a reduction of competition for pollinators in species with "original" traits and an increase of competition in species with a large number of interaction partners [@gibson_can_2012; @carvalheiro_potential_2014]. 
Alternatively, it might also be the case that abundance (for example, in terms of flower or pollen counts) is the dominant force driving pollen transfer [@seifan_outcome_2014]. 
Abundant plant species might experience a dilution of available pollinators [@feinsinger_effects_1987; @feldman_when_2004] but might also receive more effective visits by capitalising on a larger share of both visits and the pollen carried by pollinators [@stavert_plant_2019].
In this case, a potential reduction in the absolute amount of conspecific pollen received could be compensated by an increase in the amount of conspecific pollen relative to heterospecific pollen.
Altogether, it is clear that these ecological factors can indeed shape pollen deposition at the community level. 
However, we still do not understand their relative importance and **the trade-offs** that might exist between them.

Here, we investigate pollen deposition dynamics at the community scale using empirical data from eleven plant-pollinator communities in the Argentinian Pampas. 
First, we investigate the relative contribution that four ecological factors have to the pollination service. 
Specifically, we hypothesise that there are trade-offs on how these factors affect the quantity and purity of conspecific pollen deposition. 
While quantity and purity should decrease for plants that share many pollination partners, other factors like the plant's functional originality, its relative floral abundance, and its visitation patterns should have the potential to compensate for this decrease
Second, we examine how these four factors that might affect pollen deposition can change across communities where species are present. 
Because these factors may affect the pollination service in contrasting ways, and a species role is relative to other species in the community, we predict that species present in multiple communities should be flexible enough to compete for pollinators under different community contexts.

# Methods

We collected data from eleven co-flowering plant communities and their pollinators in three locations in the Argentinian Pampas. 
In each location, we sampled two restored and two agricultural fragments, except in one located in the Flooding Pampas, where we were only able to sample one restored fragment due to the lack of available sites.

## Factors affecting quantity and purity of pollination service

Our first objective was to investigate the relative contribution that different ecological factors have on pollen deposition. 
Generally speaking, we expect that any factor that increases the amount of conspecific pollen deposited in stigmas, both in quantity and purity relative to heterospecific pollen, also has a positive effect on the pollination service.
Specifically, we investigated the effect of (*i*) a plant's number of shared pollinator species, (*ii*) a plant's abundance relative to the rest of the community, (*iii*) the mean visit potential---a metric that combines the amount and type of pollen carried by floral visitors and the number of visits it receives from them, and (*iv*) the plant's functional originality [@laliberte_distance-based_2010]. 
See *Data Analysis* section below for more details on these four factors.

### Data collection

In each of the studied communities, we quantified pollen deposition in a subset of plant species between December 2010 and February 2011. 
This subset comprised between three and nine common insect-pollinated (entomophilous) plant species that were flowering during the sampling period. 
Based on data from previous years [@marrero_effect_2014], we chose plant species such that they cover a wide range on a specialization-generalization gradient as well as a wide range of abundances. 
In each of the selected plants, we removed all flowers leaving only buds that were expected to go into florescence on the next day. 
Two days after flowering, we collected all remaining flowers and counted the number of conspecific and heterospecific pollen grains in their pistils. 
More details can be found in @marrero_agricultural_2016.

To obtain the number of shared pollinators for each species, we collected data to construct qualitative and quantitative pollination networks. 
Qualitative networks were constructed based on ten-hour observations of floral visits in each fragment. 
Quantitative networks were constructed using two 50 m randomly located transects in each fragment. 
We counted and collected all floral visitors found in a 2 m wide strip while walking at a pace of 10 m per minute [@marrero_effect_2014; @memmott_structure_1999]. 
We visited the transects each month between November 2010 and March 2011. 
To obtain floral abundance, we counted all units of floral attraction found during an independent sampling of the same transects used to construct the quantitative visitation networks. 
To estimate visit potential, we need to construct pollen transfer networks in addition to the visitation networks. 
To do this, we examined the pollen loads present on the floral visitors collected [@marrero_exotic_2017]. 
When the pollen count on an individual animal was estimated to be less than 2,000 grains, we identified every grain to the species level when possible and to pollen complexes when it was not. 
When the pollen count was above 2,000 grains, we classified approximately 50% of pollen and total pollen counts were extrapolated [@bosch_plant-pollinator_2009]. 
Finally, we also recorded morphological traits that relate to plant type (herb, shrub, climber), life cycle (annual, perennial), flower colouration, phenology, and whether the species is native in the study region. 
More details can be found in Marrero *et al.* (2014 and 2017).

### Data analysis

To investigate the impact of ecological factors on pollination services, we used two sets of linear mixed models (LMM) with bootstrap resampling.
The response variables for these model sets were the number of conspecific and heterospecific pollen grains deposited per stigma in flowers open to animal-mediated pollination. 
We used LMMs in which pollen loads were log-transformed because these models offered a better fit than equivalent GLMMs with Poisson (or quasi-Poisson) error structure. 
Models were fitted using the R package `nlme 3.1-131` [@pinheiro_nlme_2018].

Because the amount of deposited pollen can vary widely across species, and potentially also across communities, we evaluated two possible structures for the random effects: one that includes a random intercept for plant species, and one that treats measures from species across different communities independently. 
We selected the best random structure by comparing the median Akaike Information Criterion for small samples (AICc).

As fixed predictors in the models, we included the four ecological factors described above. 
Specifically, we calculated the number of shared pollinators for each plant species by pooling data from the qualitative and quantitative pollination networks. 
To calculate the plants' relative floral abundance in their community, we aggregated floral counts for each species. 
We then calculated the mean visit potential of pollinator species *i* to plant species *j* as \[o_{ij} = \frac{v_{ij}}{v_i} \frac{p_{ji}}{p_j}\], where *v~ij~* is the observed number of visits by *i* to *j*, *p~ji~* is the number of pollen grains from *j* attached to *i*, *v~i~* is the total number of visits performed by *i*, and *p~j~* is the total number of grains carried by *j*.
We log-transformed the number of shared pollinators, floral abundance, and visit potential before including them in the model.

Finally, functional originality is defined as the distance of a species from the community trait average--the centroid of functional space of the community [@laliberte_distance-based_2010; @coux_linking_2016]. 
To include phenological variation, we treated floral abundance in each of the survey months (November to March) as a "trait" in our analysis. 
To account for the non-independence of floral counts and weight all traits equally, we assigned a weight of 1/5 to these abundances (one for each month). 
We scaled all traits prior to calculating the centroid of the functional space and calculated the species-specific functional coordinates using the R package `FD 1.0-12` [@laliberte_fd_2014]. 
Finally, we scaled all four factors to have a zero mean and unitary variance.

To estimate the coefficients, perform model selection, and quantify the associated uncertainty, we used a combination of multi-model inference and bootstrap resampling with 99 replicates. 
First, we performed model selection using the AICc and determined the likelihood of each candidate model (a particular combination of predictors) by calculating the median \(\Delta\)AICc (relative to the most likely model) for each bootstrap sample. 
As we wanted model coefficients from more likely candidate models to carry more weight in our results, we sampled the coefficients for our factors proportionally to the likelihood of their candidate model. 
Finally, we used these distributions of the model coefficients to estimate their mean impact on the pollination service (in terms of quantity and purity of conspecific pollen deposition).

## Flexibility of plant strategies

Our second objective was to tease apart whether and how these factors that might affect pollen deposition might change across communities species are present.
If community context plays a relatively small role, or alternatively, species are inflexible in regards of these factors, we would expect plants of the same species to fill similar pollination niches across different communities. 
Alternatively, if the community plays a large role and plant species are flexible, we should be able to observe differences in the realised niche a plant species occupies across communities. 
To test this, we first used a principal component analysis (PCA) of the four ecological factors (number of shared pollinators, floral abundance, visit potential, and trait originality). 
We scaled factors across the whole study to ensure that the PCA space does not change according to the species present in each community. 
We define a species' niche in a community as its coordinates in PCA space. 
For each species that was present in two or more communities, we then calculated (*i*) the median distance between the strategies that the species uses in different communities and (*ii*) the area of the convex hull defined by these points in the first two principal components (only for species present in three or more communities). 
We then compared the magnitude of these two metrics to those obtained with 99 Monte Carlo randomizations in which we replaced the strategy of the focal plant species by that of another randomly selected species in the dataset.

# Results

## Factors affecting quantity and purity of pollination service

```{r model-fit-statistics}
drake::loadd(r2_values)
```

We then examined the potential roles played in pollen deposition by four ecological factors (number of shared pollinators, abundance, mean visit potential, and functional originality). 
We found that our models of pollen deposition had high explanatory power (the coefficient of determination R^2^ ranged between `r min(r2_values$r2c_min)` and `r max(r2_values$r2c_max)`) although a large portion of the explanatory power came from the random effects (Table \@ref(S-tab:table-R2)).
As determined by AICc, the random structure best supported by the data was the one that fit an intercept for each species in each community (as opposed to a common intercept for each species irrespective of the community to which they belong). 
This structure was best for both the models of conspecific and heterospecific pollen (Table \@ref(S-tab:table-aic-random-models)).

Of the four factors we considered, we found that a plant's mean visit potential and relative floral abundance were the most important at predicting pollen deposition in plant stigmas (Fig. \@ref(fig:fig-average-qual-quant)a). 
Surprisingly, the number of shared pollinators was comparatively unimportant, particularly for models of heterospecific pollen deposition, as it was only ever included in models with relatively large AICc values (Table \@ref(S-tab:table-aic-fixed-models)).

```{r fig-average-qual-quant, fig.height= 2.28*2, fig.width=width('single'), out.height="4.56in", out.width="3.57in", out.extra="center", fig.cap="Effect of ecological factors on the pollination service. (a) The plant's visit potential and relative floral abundance are the most important factors determining the deposition of conspecific and heterospecific pollen. Meanwhile, the number of shared pollinators was generally less important. The graph shows the relative importance calculated as the sum of the Akaike weights of the candidate models that included the selected factor. (b) The association between ecological factors and heterospecific pollen tended to align with their association with conspecific pollen. Visit potential and functional originality had a positive association with pollen deposition, while floral abundance and the number of shared pollinators had a negative association. The plot shows the distribution of the effects (across 99 bootstrap replicates) of the four ecological factors for conspecific and heterospecific pollen. (c) The end result of these associations is that only the plants' functional originality has a positive impact on both the quantity and purity of conspecific pollen deposition (relative to heterospecific pollen). The plot shows the model averaged mean effect ($\\pm$ SE of 99 bootstrap replicates)."}
height1 <- 1
height2 <- 1.28
# pdf(width = 3.57 , height = sum(c(height1, height2))*2)
# fig_coef_avg <- plot_coefficient_averages(coefficient_averages, variable_importance)
# width_imps <- 6.5/5*2
# mult <- 2.455 # for 2/3 is xx, for 3/5 is 2.355
# width_coef_1 <- (6.5 - width_imp) / (1 + mult)
# width_coef_2 <- 5.6 - width_imp - width_coef_1
# fig_var_importance <- plot_variable_importance(variable_importance)
drake::loadd(fig_var_importance)
drake::loadd(fig_coef_avg)
drake::loadd(fig_coefficient_averages)

var_impo <- cowplot::plot_grid(
  fig_var_importance[[1]],
  fig_var_importance[[3]],
  fig_var_importance[[2]],
  ncol = 3,
  rel_widths = c(1,0.75,1),
  align = "h"
)

bottom <- cowplot::plot_grid(
  fig_coef_avg[[1]],
  fig_coef_avg[[3]],
  fig_coef_avg[[2]],
  rel_widths = c(1,0.75,1), 
  ncol = 3,
  align = "h"
)

cowplot::plot_grid(
  var_impo, 
  fig_coefficient_averages,
  bottom, 
  ncol = 1, 
  rel_heights = c(height1, height1 + height2, height2)
)

# dev.off()
```

We found that the relationship between each of the ecological factors and pollen deposition was similar for both conspecific and heterospecific pollen. 
That is, strategies that were associated with an increase in conspecific pollen were also associated with an increase in heterospecific pollen deposition. 
Specifically, the plants' mean visit potential had a positive effect on pollen deposition (Fig. \@ref(fig:fig-average-qual-quant)b). 
However, the effect size was slightly larger for heterospecific than for conspecific pollen. 
This indicates that, although there is a positive association between visit potential and the quantity of pollen deposition, there is a negative relationship with its purity (Fig. \@ref(fig:fig-average-qual-quant)c). 
In contrast, a plants' relative floral abundance negatively affected its deposition quantity, but the mean difference between the coefficients in the models indicates a positive association with purity (Fig. \@ref(fig:fig-average-qual-quant)c). 
The third most important factor, functional originality, had a positive, although comparatively smaller, association with both the quantity and purity. 
Finally, the number of shared pollinators had negative and neutral associations with conspecific and heterospecific pollen, respectively, but these impacts were small when compared to the other factors. 
Although the ecological factors were positively correlated (Fig. \@ref(S-fig:fig-correlation)), the collinearity between predictors did not qualitatively affect our findings (Fig. \@ref(S-fig:fig-model-results)).

## Flexibility of plant strategies

```{r pca-stats}
drake::loadd(pcas)
drake::loadd(permanova_plant_distances)
chosen_threshold <- 0
this_pca <- pcas %>%
    purrr::keep(~ .$call$X$pca_type[1] == "across") %>%
    purrr::keep(~ .$call$X$threshold[1] == chosen_threshold) %>%
    extract2(1)
median_distances <- permanova_plant_distances %>% 
  dplyr::filter(na_threshold == 0, metric == "median_dist", !is.na(value)) %$%
  value
```

We used a PCA of the ecological factors--species matrix to investigate whether plants' "strategy" towards pollen deposition is similar across communities or whether they are flexible and therefore a reflection of the community context. 
The first two PCA components explained `r round(this_pca$eig[2, 3])`% of the total variance (Fig. \@ref(fig:fig-pca)a). 
The first component was dominated by visit potential and relative abundance while the second component was dominated by the number of shared pollinators and the plant's functional originality. 
When we locate the species that were sampled in more than one community in the first two PCA components (Fig. \@ref(fig:fig-pca)b), we observe that the niches of any given species do not tend to be close to each other. 
Indeed, when we measured the median distance between the plants' coordinates, we found that it was only significantly smaller than that of randomisations for only `r numbers2words(sum(median_distances < 0.05))` of the `r numbers2words(length(median_distances))` species analysed (Fig. \@ref(fig:fig-pca)c). 
```{r fig-pca, fig.width=3.57, fig.height=6.65, out.width="3.57in", out.height="6.65in",  out.extra="center", fig.cap="The flexibility of plant strategies. (a) The two first components explain a large proportion of the total variance. (b) When plants that were sampled in more than one community are plotted in terms of these two components, we observe that their points, which represent the strategy (the particular combination of ecological factors) of that species in its community, do not seem to be grouped by plant species. (c) This was confirmed using Monte Carlo randomizations of the median distance between strategies of a plant species. Only two of the examined species had strategies that were less flexible than would be expected at random."}
height0 <- 1.55
height1 <- 2.7
height2 <- 2.4
strategy_fig_height <- sum(height0, height1, height2)
# fig_pca <- plot_pca(pcas, chosen_threshold = 0)
# fig_distances <- plot_permanova_dist(permanova_plant_distances, permanova_site_distances)
# fig_pca_contrib <- plot_pca_variances_and_contributions(pcas, chosen_threshold = 0)
# pdf(width = 3.57, height = strategy_fig_height)
drake::loadd(fig_pca_contrib)
drake::loadd(fig_pca)
drake::loadd(fig_distances)
cowplot::plot_grid(
  fig_pca_contrib,
  fig_pca,
  fig_distances,
  ncol = 1, 
  rel_heights = c(height0, height1, height2), 
  align = "v"
)
# dev.off()
```

# Discussion

Our results suggest that community context plays a central role in determining the net cost or benefit of sharing pollinators.
<!-- DBS: In the intro, it's about net outcome. Here, it's about net cost or benefit. Which is it? -->
First, we found that pollinator sharing can lead to both net facilitative and net competitive effects to plant species in a community. 
Second, we found that multiple ecological factors can modulate the intensity of competition; however, conspecific and heterospecific pollen deposition are tightly coupled and this creates a clear trade-off between the quantity and purity of pollination [@thomson_effects_2019]. 
Third, we found that the way these factors influence competition is strongly shaped by the community context. 
That is, if these factors collectively define a species' pollination niche,
<!-- DBS: Yet another way to phrase this concept... -->
the same plant species can occupy dramatically different niches depending on the community to which it belongs.

The criterion we used to identify the effect of animal-mediated pollination suggested that neither facilitation nor competition are dominant in pollination communities.
A previous study showed that, in diverse pollination communities, sharing pollinators does not necessarily translate to net competition [@tur_evaluating_2016]. 
The overall extent of facilitation/competition of pollination communities can depend on factors like the adversity of the environment [@tur_evaluating_2016; @callaway_positive_2002] or how disturbances affect pollinator populations [@stavert_exotic_2017]. 
However, why some species experience more competition than others was not understood. 
Although we were unable to measure the fitness outcomes, here we show that the coexistence of facilitative and competitive effects of animal-mediated pollination can be at least partially explained by trade-offs in the pollination service--both the short and the long term. 
These trade-offs arise when plants simultaneously maximise the deposition of conspecific pollen and minimise that of heterospecific pollen.

In the short term, being a specialist and sharing no pollinators might reduce competition [@muchhala_competition_2010] and hence be preferable, not only due to costs to male fitness [@muchhala_interspecific_2012; @morales_interspecific_2008], but also because, as we show here, sharing pollinators reduces both the quantity and purity of the conspecific pollen deposited. 
However, over long periods of time, there could be a risk associated with a specialist plant having few pollinators [@ricketts_tropical_2004]. 
To ensure long-term survival, it is thus likely that plants also need to balance this risk with the costs of sharing pollinators [@aizen_specialization_2012]. 
One possible solution is to share pollinators *and* have original traits---as we show that trait originality is generally beneficial to pollen deposition and it is commonly thought that species that are further from others in trait space benefit from reduced competition. 
Yet, there are two possible caveats to this strategy that highlight the interrelatedness of the ecological factors. 
First, in a mutualism context, it is also possible that trait originality could come at the cost of being less ‘apparent’ to pollinators [@reverte_pollinators_2016]. 
Second, the negative relationship between originality and generalism  [@carvalheiro_potential_2014] has been shown to depend on plant abundance [@coux_linking_2016], with generalist species being able to also have original traits only when they are abundant enough to provide a valuable reward to make visiting worthwhile to pollinators.

Visit potential (high pollen and visits) and floral abundance, which were the most important predictors of pollen deposition here, introduced an even more explicit trade-off between gaining conspecific pollen and avoiding heterospecific pollen. 
Receiving high visitation increases conspecific pollen deposition but increases heterospecific pollen deposition to a greater extent---even when the visitors are likely to carry a high proportion of conspecific pollen [@fang_paradoxical_2016]. 
Contrastingly, being abundant reduces the amount of conspecific pollen deposited and simultaneously reduces heterospecific pollen at a faster rate. 
Our results corroborate the importance that two-species studies have ascribed to visitation and abundance [@feldman_when_2004; @munoz_presence_2008; @morales_interspecific_2008], 
but they also suggest that [because visitation, pollen production and abundance are usually correlated; @sargent_role_2006] balancing the pros and cons of sharing pollinators at the community level is not trivial and might even be partially responsible for the diversity of plant-pollinator communities [@benadi_frequency_2018].

We observed, as expected, that the effects of pollen deposition can vary widely among species. 
For instance, while the fitness of some plant species can be hurt even by low amounts of heterospecific pollen, for others, fitness can instead be limited by the amount of conspecific pollen [@campbell_mechanism_1985; @arceo-gomez_interactive_2019]. 
Alternatively, plant species can also differ substantially in the extent to which self- vs. outcross-pollen differ in their value for fertilization. 
The difference can be particularly relevant for species that are not self-fertile or those in which self-fertilization is rarely effective due to a temporary separation in the maturation of the sexes (dichogamy).

Importantly, we show here that the balances between costs and benefits are determined not only by species identity but also by the community to which plants belong. 
Specifically, most plant species appear to be flexible enough to adopt markedly different niches in different communities. 
<!-- Although there are many exceptions, some plant families (Asteraceae for example) are often generalists in their communities while others (Orchidaceae) are known to be more commonly specialised [@johnson_generalization_2000].  -->
<!-- DBS: I'd cut the sentence above. -->
From an evolutionary perspective, our results suggest that selection for a particular strategy might say something about the community in which a species has typically inhabited during its evolutionary history. 
Furthermore, from a more applied perspective, flowering plants are sometimes introduced to attract pollinators on other nearby plants. 
<!-- DBS: Below, we do the 'on the one hand', 'on the other (hand)' dance for the third or fourth time. We should change some of them to spice things up. -->
On the one hand, our results suggest that introduced plants that increase the relative originality of natives (e.g. through distinct phenology) might have positive effects [@gibson_can_2012]. 
On the other, because different niches can lead to different outcomes across communities, our results also highlight the difficulties involved in predicting whether the introduced plant species will facilitate or compete with neighbours [@bartomeus_contrasting_2008]. 
Other factors that we were unable to measure (e.g. pollinator behaviour and densities or the spatial context) have also been shown to play a role in the outcome of animal-mediated pollination [@ye_competition_2014; @flanagan_effects_2011; @cariveau_spatially_2009; @thomson_effects_2019].
Nevertheless, our results indicate that the strategies a plant might use to successfully minimise competition for pollination (or maximise facilitation) must be determined relative to other species in the community, rather than an absolute property of the species itself.

Overall, pollination communities might not necessarily be the paradigmatic mutualistic communities we often believe. 
Instead, we confirm that plant competition is pervasive in pollination communities. 
The potential strategies to minimise competition are likely to involve strong trade-offs in the pollination service both in the short and long-term. 
<!-- Specifically, although not sharing pollinators and being functionally unique reduces competition for pollination by increasing the quantity and purity of conspecific pollen deposition.  --> 
<!-- Remarkably, this increase is surprisingly small when compared to the effect of relative abundance and visitation.  -->
<!-- Interestingly, these two factors also involve the largest trade-offs between increasing deposition of conspecific pollen and reducing that of heterospecific pollen.  -->
Most remarkably, despite large differences in the quantity and purity of conspecific pollen deposition among species, our results provide solid evidence that competition for pollination is inherently a community process and the same species can occupy different niches in different communities. 
<!-- Overall, species coexistence is not solely determined by direct and indirect mutualistic relationships, but also strongly shaped by the multiple intertwined ecological factors that shape competition for pollination at the community level.  -->
Many of the widely used theoretical models of plant-pollinator communities do not account for the negative effects of sharing pollinators [but see @rohr_structural_2014 and similar]\R{theoretical-models}. 
However, achieving a better understanding of species coexistence and how plant biodiversity is supported by pollination communities will require seeing them as both mutualistic and competitive communities [@johnson_coexistence_2019].

<!-- DBS: If we're really looking for a paragraph to cut, we can cut this last one in its entirety. -->

<!-- ESC: hi fernando. i'm edine. i'm here with my papa. have a nice day!!!!!! -->

# Acknowledgements

`r paste(drake::readd(acknowledgements), collapse = "")`

# Funding

Fieldwork was supported by grants PICT 08–12504 and 0851. 
EFC acknowledges the support from the University of Canterbury Doctoral Scholarship and a New Zealand International Doctoral Research Scholarship administered by New Zealand Education. 
DBS and JMT acknowledge the support of Rutherford Discovery Fellowships (RDF-13-UOC-003 and RDF-UOC-1002) and the Marsden Fund Council (UOC-1705), administered by the Royal Society of New Zealand Te Apārangi.

# References