### 2019-11-24 paper/draft-info.Rmd
Update draft info

---
title: "Draft information"
subtitle: "version: draft-8.0"
date: "2019-11-24"
# author: "Fernando Cagua, Hugo Marrero, Jason Tylianakis, Daniel Stouffer"
# bibliography: bibliography.bib
# csl: nature.csl
output:
  bookdown::pdf_document2:
    # base_format: rticles::peerj_article
    keep_tex: yes
    number_sections: false
toc: false
classoption: a4paper
geometry: 
  - textwidth=33em
  - textheight=48\baselineskip
# classoption: twocolumn
header-includes:
  - \usepackage{booktabs}
  - \usepackage{setspace}
  - \usepackage{lineno}
  - \usepackage{xr}
  - \usepackage[utf8]{inputenc}
  - \newcommand{\R}[1]{\label{#1}\linelabel{#1}}
  - \newcommand{\lr}[1]{page~\pageref{#1}, line~\lineref{#1}}
  - \externaldocument[M-]{manuscript}
---

\onehalfspacing

## Requests to co-authors

* Title suggestions/feedback?
* Feedback primarly on abstract, introduction and discussion prior to submission.

## Changes since the last draft

* Adjusted langage away from competition for pollination towards pollination service and pollen deposition across the paper.
* Removed the section where we calculated competition/facilitation in ecological communities
* Other minor adjustments to address comments by reviews received when submitted to Proceedings B
* Rewrote abstract
* Updated title